Something about Performance, 2D Component of NEMO. (TBC)

### 2015-03-03 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
Beginning to add OpenMP results and description

Our aim in this work is to achieve portable performance. Consequently,
we have performed tests with the hardware and compilers listed in
Table~\ref{TABLE_compilers}. Where a compiler is available on a given
piece of hardware, the version number used in this work is specified.
The Intel Haswell-based system (Xeon E5-1620 v2) has a clock speed of
3.7~GHz and 10~MB of last-level cache. The Intel Ivy Bridge-based
system (Xeon E5-2697) has a clock speed of 2.7~GHz and a last-level
cache of 30~MB. The CPU in the IBM BG/Q system is built around the
PowerPC A2 core with a clock speed of 1.6~GHz. The chip has 32~MB of
last-level cache.

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
\caption{The matrix of compilers and CPUs used in this work. The use
  of a compiler on a given CPU is indicated by the specification of
  the version of the compiler in the relevent element. No entry
  indicates that a compiler was not available/used on that CPU.}
\label{TABLE_compilers}
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
                 & \multicolumn{4}{c|}{Compiler}             \\
\hline
                 & Gnu   & Intel       & Cray    & IBM bgxlf\\
\hline
Intel Haswell    & 4.8.2 & 14.0.0      &         &          \\
Intel Ivy Bridge & 4.9.1 & 15.0.0.090  & 8.3.3   &          \\
IBM A2 (BGQ)     &       &             &         &   --     \\
IBM Power 8      &       &             &         &          \\
\hline
\end{tabular}
\end{table}

@@

Before applying any code transformations, we benchmark the original
version of the code. We also benchmark the vanilla, unoptimised
version after it has been re-structured following the \psykal
approach. In addition to this benchmarking, we profile these versions
of the code at the algorithm level (using a high-level timing
API). The resulting profiles are given in
Table~\ref{TABLE_profile}. The Momentum section dominates the profile
of both versions of the code, accounting for around 70--80\% of the
wall-clock time spent doing time-stepping. It is also the key section
when we consider the performance loss when moving from the original to
the \psykal version of the code; it slows down by a factor of
two. Although less significant in terms of absolute time, the
Continuity section is also dramatically slower in the \psykal version,
this time by a factor of three. In contrast, the performance of the
Time-update and Boundary Condition regions are not significantly
affected by the move so the \psykal version.

@@

\subsection{Addition of `safe-address' directives}
\label{sec_safe_address}

@@

\subsection{In-line Momentum kernel bodies into middle-layer code}
\label{sec_inline_mom}

@@

\subsection{Force SIMD vectorisation of the Momentum kernels using directives}

@@

\subsection{Work around limitations related to derived types}

@@

\subsection{In-line the Continuity kernel}
\label{sec_cont_inline}

@@

\subsection{In-line remaining kernels (BCs and time-update)}
\label{sec_inline_all_kernels}

@@

\subsection{In-line field-copy operations}
\label{sec_inline_fldcopy}

@@

@@

>>>>

Our aim in this work is to achieve portable performance. Consequently,
we have performed tests with the hardware and compilers listed in
Table~\ref{TABLE_compilers}. Where a compiler is available on a given
piece of hardware, the version number used in this work is specified.
The Intel Haswell-based system (Xeon E5-1620 v2) has a clock speed of
3.7~GHz and 10~MB of last-level cache. The Intel Ivy Bridge-based
system (Xeon E5-2697) has a clock speed of 2.7~GHz and a last-level
cache of 30~MB.

We first describe the code transformations performed for the serial
version of nemolite2d.  We then move on to the construction of
parallel versions of the code using OpenMP and OpenACC. Again, we
describe the key steps we have taken in this process in order to
maximise the performance of the code.

\subsection{Transformations of Serial Nemolite2D}

% An example of a floating table. Note that, for IEEE style tables, the 
% \caption command should come BEFORE the table. Table text will default to
% \footnotesize as IEEE normally uses this smaller font for tables.
% The \label must come after \caption as always.
%
\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
\caption{The matrix of compilers and CPUs used in this work. The use
  of a compiler on a given CPU is indicated by the specification of
  the version of the compiler in the relevent element. No entry
  indicates that a compiler was not available/used on that CPU.}
\label{TABLE_compilers}
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
                 & \multicolumn{4}{c|}{Compiler}            \\
\hline
                 & Gnu   & Intel       & Cray    & IBM      \\
\hline
Intel Haswell    & 4.8.2 & 14.0.0      &         &          \\
Intel Ivy Bridge & 4.9.1 & 15.0.0.090  & 8.3.3   &          \\
IBM Power 8      &       &             &         &          \\
\hline
\end{tabular}
\end{table}

@@

Before applying any code transformations, we first benchmark the
original, serial version of the code. We also benchmark the vanilla,
unoptimised version after it has been re-structured following the
\psykal approach. In addition to this benchmarking, we profile these
versions of the code at the algorithm level (using a high-level timing
API). The resulting profiles are given in
Table~\ref{TABLE_profile}. The Momentum section dominates the profile
of both versions of the code, accounting for around 70--80\% of the
wall-clock time spent doing time-stepping. It is also the key section
when we consider the performance loss when moving from the original to
the \psykal version of the code; it slows down by a factor of
two. Although less significant in terms of absolute time, the
Continuity section is also dramatically slower in the \psykal version,
this time by a factor of three. In contrast, the performance of the
Time-update and Boundary Condition regions are not significantly
affected by the move to the \psykal version.

@@

\subsubsection{Addition of `safe-address' directives}
\label{sec_safe_address}

@@

\subsubsection{In-line Momentum kernel bodies into middle-layer code}
\label{sec_inline_mom}

@@

\subsubsection{Force SIMD vectorisation of the Momentum kernels using directives}

@@

\subsubsection{Work around limitations related to derived types}

@@

\subsubsection{In-line the Continuity kernel}
\label{sec_cont_inline}

@@

\subsubsection{In-line remaining kernels (BCs and time-update)}
\label{sec_inline_all_kernels}

@@

\subsubsection{In-line field-copy operations}
\label{sec_inline_fldcopy}

@@

\subsection{Construction of OpenMP-Parallel Nemolite2D}

For this part of the work we began with the optimal \psykal version of
the code, as obtained after applying the various transformations
described in the previous section. As with the transformations of the
serial code, our purpose here is to determine the functionality
required of the PSyclone code generation system.

@@

\subsubsection{OpenMP Version}

For the OpenMP-parallel version of nemolite2d we began with the
optimised form of the serial version, as described in the previous
section.

\begin{figure}
\centering
\includegraphics[width=2.8in]{omp_scaling_256_gnu}
\caption{Performance of the OpenMP-parallel version of \psykal
  nemolite2d for the Gnu compiler on a single Intel Ivy Bridge
  socket. The 24-thread runs employed hyperthreading.}
\label{FIG_gnu_omp_256}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.8in]{omp_scaling_256_intel}
\caption{Performance of the OpenMP-parallel version of \psykal
  nemolite2d for the Intel compiler on a single Intel Ivy Bridge
  socket. The 24-thread runs employed hyperthreading.}
\label{FIG_intel_omp_256}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=2.8in]{omp_scaling_256_cray}
\caption{Performance of the OpenMP-parallel version of \psykal
  nemolite2d for the Cray compiler on a single Intel Ivy Bridge
  socket. The 24-thread runs employed hyperthreading.}
\label{FIG_cray_omp_256}
\end{figure}

\subsubsection{OpenACC Version}
